### 2019-09-04 initial/notes.tex
stash

We zoom in on Cassini State $4$, which has $\theta_4 =
-\frac{\pi}{2} + \frac{\eta \cos I}{1 - \eta \sin I}, \mu_4 = \frac{\eta \cos
I}{1 - \eta \sin I}, \phi_4 = 0$. Then, using
equations of motion
\begin{align}
    \pd{\phi}{t} &= \mu - \eta\p*{\cos I + \sin I \frac{\mu}{\sqrt{1 - \mu^2}}
        \cos \phi} ,\\
    \pd{\mu}{t} &= -\eta \sin I \sin \phi + \s*{\epsilon \p*{1 - \mu^2}},
\end{align}
we can perturbatively require $\pd{\theta}{t} = 0$ for $\epsilon \neq 0$. This
corresponds to $\eta \sin I \sin \p*{\phi_4 + \delta \phi} \approx \epsilon$, or
$\delta \phi_4 = +\frac{\epsilon}{\eta \sin I}$. This is in agreement with
Dong's result.

@@

To make the equations more anemable to analytic analysis (not simulation), let's
write down the EOM in $(\mu, \phi)$ coordinates again. The $\phi$ EOM does not
change from the tide-free case, so we can reuse earlier equation:
\begin{align}
    \pd{\phi}{\tau} &= \frac{s}{s_c}\mu
        - \p*{\cos I + \sin I \frac{\mu}{\sqrt{1 - \mu^2}} \cos \phi} ,\\
    \pd{\mu}{\tau} &= -\sin I \sin \phi +
        \epsilon \frac{2\Omega_1}{s}\p*{1 - \mu^2}
            \p*{1 - \frac{s}{2\Omega_1}\mu},\\
    \rd{s}{\tau}
        &= \epsilon 2\Omega_1 \p*{\mu - \frac{s}{2\Omega_1}\p*{1 + \mu^2}}.
\end{align}
Assuming $s \gg s_c$ the strong spin-orbit coupling regime, let's first try
assuming $\mu$ is roughly constant over the course of a precession period, then
we can average out the $\phi$ dependencies. Then $\phi$ drops out of the EOM,
and we have approximate averaged equations
\begin{subequations}\label{se:prec_avg}
    \begin{align}
        \pd{\mu}{(\epsilon\tau)} &\approx \frac{2\Omega_1}{s}\p*{1 - \mu^2}
                \p*{1 - \frac{s}{2\Omega_1}\mu},\nonumber\\
            &\approx \p*{1 - \mu^2}
                \p*{\frac{2\Omega_1}{s} - \mu},\\
        \frac{1}{\Omega_1}\rd{s}{(\epsilon\tau)}
            &\approx 2\mu - \frac{s}{\Omega_1}\p*{1 + \mu^2}.
    \end{align}
\end{subequations}
These EOM produce roughly the phase portrait \autoref{fig:phi_avg}.
\begin{figure}[t]
    \centering
    \begin{tikzpicture}
        \draw[<->] (0, 0) -- (8, 0);
        \draw[<->] (0, 0) -- (0, 5);
        \node[below right] at (8, 0) {$s$};
        \node[above left] at (0, 5) {$\mu$};
        \draw (0, 4) -- (2, 4) -- (4, 4);
        \draw[dashed] (4, 4) -- (8, 4);
        \draw (0, 4) -- (2, 4);
        \draw[domain=0:2] plot({\x}, {\x * \x / 2 + 2});
        \node[left] at (0, 4) {$1$};
        \node[left] at (0, 2) {$0$};
        \node[left] at (0, 0) {$-1$};
        \node[below] at (0, 0) {$0$};
        \node[below] at (2, 0) {$\Omega_1$};
        \node[below] at (4, 0) {$2\Omega_1$};
        \draw[domain=4:8] plot({\x}, {16 / \x});
        \draw[->] (2, 2) -- (1.5, 2.5);
        \draw[->] (4, 2) -- (3.5, 2.5);
        \draw[->] (6, 0.5) -- (5.5, 1);
        \draw[->] (7, 3.5) -- (6.5, 3);
        \draw[->] (0.5, 3) -- (1, 3.5);
    \end{tikzpicture}
    \caption{Rough phase portrait of $\phi$-averaged equations. Dashed lines
    indicate unstable zeros of at least one of the EOM, while solid lines
    indicate stable zeros of at least one of the EOM\@. The zeros are $\mu = 1$,
    which becomes unstable at $s = 2\Omega_1$, $s = 2\Omega_1/\mu$ and $s' = 0$.
    The only fixed point is $\mu = 1, s = \Omega_1$.}\label{fig:phi_avg}
\end{figure}
In the last term, we note $s \gtrsim 2\Omega_1$ initially, while $\mu \leq 1$,
so we drop both the linear contribution from $\mu$ and approximate $\p*{1 +
\mu^2} \approx 1$ so that $\rd{(s/\Omega_1)}{(\epsilon\tau)} \approx
-s/\Omega_1$.

@@

>>>>

We zoom in on Cassini State $4$, which has $\theta_4 =
-\frac{\pi}{2} + \frac{\eta \cos I}{1 - \eta \sin I}, \mu_4 = \frac{\eta \cos
I}{1 - \eta \sin I}, \phi_4 = 0$. Then, using
equations of motion
\begin{align}
    \pd{\phi}{t} &= -\mu + \eta\p*{\cos I + \sin I \frac{\mu}{\sqrt{1 - \mu^2}}
        \cos \phi} ,\\
    \pd{\mu}{t} &= -\eta \sin I \sin \phi + \s*{\epsilon \p*{1 - \mu^2}},
\end{align}
we can perturbatively require $\pd{\theta}{t} = 0$ for $\epsilon \neq 0$. This
corresponds to $\eta \sin I \sin \p*{\phi_4 + \delta \phi} \approx \epsilon$, or
$\delta \phi_4 = +\frac{\epsilon}{\eta \sin I}$. This is in agreement with
Dong's result.

@@

To make the equations more anemable to analytic analysis (not simulation), let's
write down the EOM in $(\mu, \phi)$ coordinates again. The $\phi$ EOM does not
change from the tide-free case, so we can reuse earlier equation:
\begin{align}
    \pd{\phi}{\tau} &= -\frac{s}{s_c}\mu
        + \p*{\cos I + \sin I \frac{\mu}{\sqrt{1 - \mu^2}} \cos \phi} ,\\
    \pd{\mu}{\tau} &= -\sin I \sin \phi +
        \epsilon \frac{2\Omega_1}{s}\p*{1 - \mu^2}
            \p*{1 - \frac{s}{2\Omega_1}\mu},\\
    \rd{s}{\tau}
        &= \epsilon 2\Omega_1 \p*{\mu - \frac{s}{2\Omega_1}\p*{1 + \mu^2}}.
\end{align}
Assuming $s \gg s_c$ the strong spin-orbit coupling regime, let's first try
assuming $\mu$ is roughly constant over the course of a precession period, then
we can average out the $\phi$ dependencies. Then $\phi$ drops out of the EOM,
and we have approximate averaged equations
\begin{subequations}\label{se:prec_avg}
    \begin{align}
        \pd{\mu}{(\epsilon\tau)} &\approx \frac{2\Omega_1}{s}\p*{1 - \mu^2}
                \p*{1 - \frac{s}{2\Omega_1}\mu},\nonumber\\
            &\approx \p*{1 - \mu^2}
                \p*{\frac{2\Omega_1}{s} - \mu},\\
        \frac{1}{\Omega_1}\rd{s}{(\epsilon\tau)}
            &\approx 2\mu - \frac{s}{\Omega_1}\p*{1 + \mu^2}.
    \end{align}
\end{subequations}
These EOM produce roughly the phase portrait \autoref{fig:phi_avg}.
\begin{figure}[t]
    \centering
    \begin{tikzpicture}
        \draw[<->] (0, 0) -- (8, 0);
        \draw[<->] (0, 0) -- (0, 5);
        \node[below right] at (8, 0) {$s$};
        \node[above left] at (0, 5) {$\mu$};
        \draw (0, 4) -- (2, 4) -- (4, 4);
        \draw[dashed] (4, 4) -- (8, 4);
        \draw (0, 4) -- (2, 4);
        \draw[domain=0:2] plot({\x}, {\x * \x / 2 + 2});
        \node[left] at (0, 4) {$1$};
        \node[left] at (0, 2) {$0$};
        \node[left] at (0, 0) {$-1$};
        \node[below] at (0, 0) {$0$};
        \node[below] at (2, 0) {$\Omega_1$};
        \node[below] at (4, 0) {$2\Omega_1$};
        \draw[domain=4:8] plot({\x}, {16 / \x});
        \draw[->] (2, 2) -- (1.5, 2.5);
        \draw[->] (4, 2) -- (3.5, 2.5);
        \draw[->] (6, 0.5) -- (5.5, 1);
        \draw[->] (7, 3.5) -- (6.5, 3);
        \draw[->] (0.5, 3) -- (1, 3.5);
    \end{tikzpicture}
    \caption{Rough phase portrait of $\phi$-averaged equations. Dashed lines
    indicate unstable zeros of at least one of the EOM, while solid lines
    indicate stable zeros of at least one of the EOM\@. The zeros are $\mu = 1$,
    which becomes unstable at $s = 2\Omega_1$, $s = 2\Omega_1/\mu$ and $s' = 0$.
    The only fixed point is $\mu = 1, s = \Omega_1$.}\label{fig:phi_avg}
\end{figure}
In the last term, we note $s \gtrsim 2\Omega_1$ initially, while $\mu \leq 1$,
so we drop both the linear contribution from $\mu$ and approximate $\p*{1 +
\mu^2} \approx 1$ so that $\rd{(s/\Omega_1)}{(\epsilon\tau)} \approx
-s/\Omega_1$.

@@

\section{Misc}

\subsection{Canonical Coordinates}

Consider canonical coordinates $x = \sin \frac{\theta}{2}\cos \phi, y = \sin
\frac{\theta}{2}\sin \phi$, then the Hamiltonian can be written
\begin{align}
    H &= -\frac{1}{2}\cos^2\theta
        + \eta \p*{\cos \theta \cos I - \sin I \sin \theta \cos \phi}
        ,\nonumber\\
        &= -\frac{1}{2}\p*{1 - 2\sin^2\frac{\theta}{2}}^2
            + \eta\p*{\p*{1 - 2\sin^2\frac{\theta}{2}}\cos I
            - 2 \sin \frac{\theta}{2}\cos \phi \cos \frac{\theta}{2}\sin I}
            ,\nonumber\\
        &= -\frac{1}{2}\p*{1 - 2\p*{x^2 + y^2}}^2
            + \eta\p*{\p*{1 - 2\p*{x^2 + y^2}}\cos I
                - 2\sin I x\sqrt{1 - \p*{x^2 + y^2}}}.
\end{align}

\subsection{Destruction of CS via Tides}

Let's return to the simple EOM considered in the constant $\eta$ tides case,
where
\begin{align}
    \pd{\phi}{t} &= -\mu + \eta\p*{\cos I + \sin I \frac{\mu}{\sqrt{1 - \mu^2}}
        \cos \phi} ,\\
    \pd{\mu}{t} &= -\eta \sin I \sin \phi + \s*{\epsilon \p*{1 - \mu^2}},
\end{align}
We want to determine where tides destroys CS2. We saw before that tides tends to
shift CS2/CS4 towards each other. In fact, we can immediately see where they
will collide: in the $\dot{\mu}$ equation, no $\phi$ is stationary if
\begin{equation}
    \epsilon > \frac{\eta \sin I}{1 - \mu^2}.
\end{equation}
We can substitute in $\mu_2, \mu_4 \sim \eta \cos I$ to obtain
\begin{equation}
    \epsilon_{c} \approx \frac{\eta \sin I}{1 - \eta^2\cos^2 I}.
\end{equation}

Of course, we can also solve perturbatively to higher order for the corrections
to CS2/CS4 about their $\epsilon = 0$ values; we have already found before the
first order correction to $\phi$
\begin{equation}
    \phi_{4, 2} = \z*{0, \pi} \pm \frac{\epsilon\p*{1 - \mu^2}}{\eta \sin I}
        \approx \z*{0, \pi} \pm \frac{\epsilon}{\eta \sin I} +
        \mathcal{O}\p*{\epsilon^2}.
\end{equation}
This can then be substituted back in to find the correction to $\mu_{4, 2}$
\begin{equation}
    \mu_{4, 2} \approx \frac{\eta \cos I}{1 \mp \eta \sin I}
        \mp \frac{1}{2}\frac{\eta^2 \sin I \cos I}{1 \mp \eta \sin I}
            \p*{\frac{\epsilon}{\eta \sin I}}^2 + \mathcal{O}\p*{\epsilon^3}.
\end{equation}
These corrections and $\epsilon_c$ can be found in good agreement w/ numerics in
\lstinline{99_misc/0_stab.png}.

Now, a slightly more subtle question: is it possible for CS2 to still exist but
lose stability? Consider that for some $\epsilon\lesssim \epsilon_c$, the
equilibria lie at $\phi_{2, 4} = \pi/2 \pm \Delta \phi$, where $\epsilon_c -
\epsilon = \eta \sin I \frac{\Delta \phi^2}{2}$, then we consider perturbation
about these equilibria $\phi_{2, 4} = \pi/2 \pm \Delta \phi + \delta \phi$. The
linearized equations then read
\begin{subequations}
    \begin{align}
        \dot{\phi} &\approx -\delta \mu \mp \eta^2 \sin I \cos I \delta \phi,\\
        \dot{\mu} &\approx \pm\eta \sin I \Delta \phi \delta \phi.
    \end{align}
\end{subequations}
The eigenvalue equation for this first-order system reads
\begin{equation}
    \lambda^2 \pm \eta^2 \sin I \cos I \pm \eta \sin I \Delta \phi = 0.
\end{equation}
This has real solutions if
\begin{equation}
    \p*{\eta^2 \sin I \cos I}^2 \mp \eta \sin I \Delta \phi > 0.
\end{equation}
If $\epsilon \sim \epsilon_c / 2$, i.e.\ not too close to critical, then
$\Delta \phi^2 \sim 1$, so except for $\epsilon_c - \epsilon \ll \epsilon_c$,
the second term dominates the first and we have one real and one imaginary root,
corresponding to the two CSs. Otherwise, we can identify that for
\begin{align*}
    \Delta \phi &\leq \eta^3 \sin I \cos^2 I,\\
    \epsilon_c - \epsilon &\leq \frac{\eta^7 \sin^3 I \cos^4 I}{2},
\end{align*}
that CS2 becomes unstable.
